---
title: "Biostat776project3"
author: "Wendy Espinoza Camejo"
date: "2023-10-14"
output: html_document
---
```{r}
library(tidyverse)
library(tidytuesdayR)
```

```{r}
library("here")
rds_files <- c("b_lyrics.RDS", "ts_lyrics.RDS", "sales.RDS")
## Check whether we have all 3 files
if (any(!file.exists(here("data", rds_files)))) {
    ## If we don't, then download the data
    b_lyrics <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv")
    ts_lyrics <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv")
    sales <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv")

    ## Then save the data objects to RDS files
    saveRDS(b_lyrics, file = here("data", "b_lyrics.RDS"))
    saveRDS(ts_lyrics, file = here("data", "ts_lyrics.RDS"))
    saveRDS(sales, file = here("data", "sales.RDS"))
}
```

```{r}
b_lyrics <- readRDS(here("data", "b_lyrics.RDS"))
ts_lyrics <- readRDS(here("data", "ts_lyrics.RDS"))
sales <- readRDS(here("data", "sales.RDS"))
```

###Part 1: Explore album sales
##Part 1A

In this section, we will do some data wrangling.

Use lubridate to create a column called released that is a Date class. However, to be able to do this, you first need to use stringr to search for pattern that matches things like this “(US)[51]” in a string like this “September 1, 2006 (US)[51]” and removes them. (Note: to get full credit, you must create the regular expression).
Use forcats to create a factor called country (Note: you may need to collapse some factor levels).
Transform the sales into a unit that is album sales in millions of dollars.
Keep only album sales from the UK, the US or the World.
Auto print your final wrangled tibble data frame.
```{r}
library(lubridate)
library(stringr)
library(forcats)
sales <- sales %>%
  mutate(released = mdy(str_trim(str_replace(string = released, pattern = "\\(..\\)\\[..\\]$", replace ="")))) 
class(sales$released)

sales <- sales %>%
 filter(!is.na(country))%>%
mutate(country = as.factor(case_when(
    country == "WW" ~ "World",
    country == "FR" ~ "FRA",
    .default = country
  )))
class(sales$country)

sales_USUKWorld <- sales %>%
  mutate(Album_sales_M = (sales / 1e6)) %>%
  filter(country == "UK"| country == "US" | country == "World")
dim(sales_USUKWorld)
print(sales_USUKWorld, n = dim(sales_USUKWorld)[1])
```

##Part 1B
Keep only album sales from the US.
Create a new column called years_since_release corresponding to the number of years since the release of each album from Beyoncé and Taylor Swift. This should be a whole number and you should round down to “14” if you get a non-whole number like “14.12” years. (Hint: you may find the interval() function from lubridate helpful here, but this not the only way to do this.)
Calculate the most recent, oldest, and the median years since albums were released for both Beyoncé and Taylor Swift.
```{r}
sales_US <- sales_USUKWorld %>%
  filter(country == "US")

sales_US_years <- sales_US %>%
  mutate(years_since_release = floor(interval(released, today())/dyears(1)))

summary_tb <- sales_US_years %>%
  group_by(artist)%>%
  summarize(Most_recent = min(years_since_release, na.rm = TRUE),
            Oldest = max(years_since_release, na.rm = TRUE),
    Median = median(years_since_release, na.rm= TRUE))
print(summary_tb)
```

##Part 1C
Calculate the total album sales for each artist and for each country (only sales from the UK, US, and World).
Using the total album sales, create a percent stacked barchart using ggplot2 of the percentage of sales of studio albums (in millions) along the y-axis for the two artists along the x-axis colored by the country.
```{r}
total_album_sales_USUKWorld <- sales_USUKWorld %>%
  group_by(artist,country) %>%
  summarize(Total_album_sales = sum(Album_sales_M))
print(total_album_sales_USUKWorld)

library(ggplot2)
graph1 <- total_album_sales_USUKWorld  %>%
  group_by(artist) %>%
  mutate(pct = Total_album_sales/ sum(Total_album_sales) * 100) %>%
  ggplot(aes(x= artist, y= pct, fill= country)) + 
  geom_bar(stat="identity") + 
   scale_y_continuous() +
    geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
    position=position_stack(vjust=0.5)) +
  theme_bw()+
  labs(title = "Percentage of album sales for Beyonce and Taylor Swift in US, UK, and the world ", subtitle = "While Beyonce has a higer record of album sales in the world, Taylor Swift is ahead in the US ", x = "Artist", y = "Percentage of sales of studio albums (in millions)", caption = "Wendy Espinoza")

print(graph1)
  
```

##Part 1D
Using the wrangled data from Part 1A, use ggplot2 to create a bar plot for the sales of studio albums (in millions) along the x-axis for each of the album titles along the y-axis.
Note:
You only need to consider the global World sales (you can ignore US and UK sales for this part).
The title of the album must be clearly readable along the y-axis.
Each bar should be colored by which artist made that album.
The bars should be ordered from albums with the most sales (top) to the least sales (bottom) (Note: you must use functions from forcats for this step).
```{r}
library(forcats)
sales_World <- sales_USUKWorld %>%
  filter(country == "World")
graph2 <- sales_World %>%
  mutate(this = fct_reorder(.f= title, .x = Album_sales_M)) %>%
  ggplot(aes(x = Album_sales_M, y= this, fill = artist)) + 
  geom_bar(stat= "identity") +
  theme_minimal()+
  labs(title = "Top Albums sales for Beyonce and Taylor Swift in the world", subtitle = "Similar record of album sales for Beyonce and Taylor Swift ", x = "Sales of studio albums (in millions)", y = "Albums", caption = "Wendy Espinoza")
  
print(graph2)
```


##Part 1E
Using the wrangled data from Part 1A, use ggplot2 to create a scatter plot of sales of studio albums (in millions) along the y-axis by the released date for each album along the x-axis.
Note:
The points should be colored by the artist.
There should be three scatter plots (one for UK, US and world sales) faceted by rows.
```{r}
graph3 <- sales_USUKWorld %>%
  ggplot(aes(x = released, y = Album_sales_M, fill = artist)) +
  geom_point(aes(colour = artist), size = 2) +
  facet_grid(country ~.) +
  theme_minimal()+
  labs(title = "Sales of studio albums in US, UK and the world", subtitle = "Taylor Swift has a higher sales of studios albums in US, UK and the world", x = "Albums released by date", y = "Sales of studio albums (in millions)", caption = "Wendy Espinoza")

print(graph3)
```

###Part 2: Exploring sentiment of lyrics
##Part 2A
Using ts_lyrics, create a new column called line with one line containing the character string for each line of Taylor Swift’s songs.
How many lines in Taylor Swift’s lyrics contain the word “hello”? For full credit, show all the rows in ts_lyrics that have “hello” in the line column and report how many rows there are in total.
How many lines in Taylor Swift’s lyrics contain the word “goodbye”? For full credit, show all the rows in ts_lyrics that have “goodbye” in the line column and report how many rows there are in total.
```{r}
library(tidyverse)
library(tidytext)
library(stringr)

class(ts_lyrics)
ts_lyrics_tb1 <-  unnest_tokens(ts_lyrics,
    output = lines,
    input = Lyrics,
    token = "lines")
dim(ts_lyrics)
dim(ts_lyrics_tb1)
#how many hello?
ts_hello <- ts_lyrics_tb1 %>%
  filter(str_detect(lines, "hello"))
ts_hello
ts_hello %>%
  count()
# how many goodbye?
ts_goodbye <- ts_lyrics_tb1 %>%
  filter(str_detect(lines, "goodbye")) %>%
  head()
ts_goodbye
ts_goodbye %>%
  count()

```

##Part 2B
Repeat the same analysis for b_lyrics as described in Part 2A.
```{r}
class(b_lyrics)
glimpse(b_lyrics)
b_lyrics_tb1 <- b_lyrics
#how many hello?
b_hello <- b_lyrics_tb1 %>%
  filter(str_detect(line, "hello"))
b_hello
b_hello %>%
  count()
#how many goodbye?
b_goodbye <- b_lyrics_tb1 %>%
  filter(str_detect(line, "goodbye"))
b_goodbye
b_goodbye %>%
  count()
```

##Part 2C
Using the b_lyrics dataset,
Tokenize each lyrical line by words.
Remove the “stopwords”.
Calculate the total number for each word in the lyrics.
Using the “bing” sentiment lexicon, add a column to the summarized data frame adding the “bing” sentiment lexicon.
Sort the rows from most frequent to least frequent words.
Only keep the top 25 most frequent words.
Auto print the wrangled tibble data frame.
Use ggplot2 to create a bar plot with the top words on the y-axis and the frequency of each word on the x-axis. Color each bar by the sentiment of each word from the “bing” sentiment lexicon. Bars should be ordered from most frequent on the top to least frequent on the bottom of the plot.
Create a word cloud of the top 25 most frequent words.
```{r}
library(dplyr)
b_lyrics_tb2 <-  b_lyrics_tb1 %>% 
  unnest_tokens(
    output = word,
    input = line) %>%
  anti_join(get_stopwords()) %>%
  count(word, sort = TRUE) %>%
 inner_join(get_sentiments("bing")) %>%
  head(n=25)
print(b_lyrics_tb2, n = dim(b_lyrics_tb2)[1])

#plot
graph4 <- b_lyrics_tb2 %>%
  mutate(this = fct_reorder(.f= word, .x = n)) %>%
  ggplot(aes(x = n, y= this, fill = sentiment)) + 
  geom_bar(stat= "identity") +
  theme_minimal()+
  labs(title = "Top 25 words used in Beyonce's songs chacterized by sentiment", subtitle = "Like and love, positive sentiment words, are the most frequently used in Beyonce's songs", x = "Frequency", y = "Words", caption = "Wendy Espinoza")
 print(graph4)

#word clouds
library(RColorBrewer)
library(wordcloud)
b_lyrics_tb2 %>%
    with(wordcloud(word, n))

```

##Part 2D
Repeat the same analysis as above in Part 2C, but for ts_lyrics.
```{r}
library(dplyr)
library(textdata)


ts_lyrics_tb2 <-  ts_lyrics_tb1 %>% 
  unnest_tokens(
    output = word,
    input = lines) %>%
  anti_join(get_stopwords()) %>%
  count(word, sort = TRUE) %>%
 inner_join(get_sentiments("bing")) %>%
  head(n=25)
print(ts_lyrics_tb2, n = dim(ts_lyrics_tb2)[1])

#plot
graph5 <- ts_lyrics_tb2 %>%
  mutate(this = fct_reorder(.f= word, .x = n)) %>%
  ggplot(aes(x = n, y= this, fill = sentiment)) + 
  geom_bar(stat= "identity") +
  theme_minimal()+
  labs(title = "Top 25 words used in Taylor Swift's songs chacterized by sentiment", subtitle = "Like and love, positive sentiment words, are the most frequently used in Taylor Swift's songs", x = "Frequency", y = "Words", caption = "Wendy Espinoza")
  
print(graph5)
### really ugly figure
#word clouds
library(RColorBrewer)
library(wordcloud)
ts_lyrics_tb2 %>%
    with(wordcloud(word, n))
```

##Part 2E
Using the ts_lyrics dataset,
Tokenize each lyrical line by words.
Remove the “stopwords”.
Calculate the total number for each word in the lyrics for each Album.
Using the “afinn” sentiment lexicon, add a column to the summarized data frame adding the “afinn” sentiment lexicon.
Calculate the average sentiment score for each Album.
Auto print the wrangled tibble data frame.
Join the wrangled data frame from Part 1A (album sales in millions) with the wrangled data frame from #6 above (average sentiment score for each album).
Using ggplot2, create a scatter plot of the average sentiment score for each album (y-axis) and the album release data along the x-axis. Make the size of each point the album sales in millions.
Add a horizontal line at y-intercept=0.
Write 2-3 sentences interpreting the plot answering the question “How has the sentiment of Taylor Swift’s albums have changed over time?”. Add a title, subtitle, and useful axis labels.

```{r}
library(textdata)
library(dplyr)
ts_lyrics_tb3 <-  ts_lyrics_tb1 %>% 
  unnest_tokens(
    output = word,
    input = lines) %>%
  anti_join(get_stopwords()) %>%
  group_by(Album, word) %>%
  count(word) %>%
 inner_join(get_sentiments("afinn")) %>%
  group_by(Album) %>%
  summarise(Average = mean(value))  
print(ts_lyrics_tb3, n = dim(ts_lyrics_tb3)[1])


sales_USUKWorld1 <- sales_USUKWorld %>%
  rename("Album" = "title")

ts_lyrics_tb4 <- ts_lyrics_tb3 %>%
  mutate(Album1 = case_when(Album == "reputation" ~ "Reputation",
                   TRUE ~ Album)) %>%
  select(-Album) %>%
  rename("Album" = "Album1")


join_data <- inner_join(sales_USUKWorld1, ts_lyrics_tb4)

graph6 <- join_data %>% 
 ggplot(aes(x = released, y = Average)) +
  geom_point(aes(size = Album_sales_M)) +
  geom_text(aes(label= Album), vjust = 1.8, hjust = 0.3) +
  geom_hline(yintercept = 0) +
  theme_minimal() +
  labs(title = "Drastic change in Taylor Swift's album sentiment and album sales", subtitle = "Taylor Swift's albums sentiment have change from positive to negative affecting her overall sales", x = "Album released dates", y = "Average sentiment", caption = "Wendy Espinoza")

print(graph6)

graph7 <- join_data %>% 
  filter(country == "US") %>%
 ggplot(aes(x = released, y = Average)) +
  geom_point(aes(size = Album_sales_M)) +
  geom_text(aes(label= Album), vjust = 1.8, hjust = 0.3) +
  geom_hline(yintercept = 0) +
  theme_minimal()+
   labs(title = "Drastic change in Taylor Swift's album sentiment and album sales", subtitle = "Taylor Swift's albums sentiment have change from positive to negative affecting her sales in the US", x = "Album released dates", y = "Average sentiment", caption = "Wendy Espinoza")
print(graph7)

```

Overtime, Taylor Swift's albums have change drastically. They went from a positive context to a negative context with a matter of 4 years. Also, her album sales have been affected by this drastic change. She went from making over 5 million per album to less than 4 millions starting in 2017.


```{r}
options(width = 120)
sessioninfo::session_info()
```

